Integrating brain and biomechanical models - a new paradigm for understanding neuro-muscular control

### 2017-11-09 papers/integration/omm_integ.tex
Adds reviewer colouring to review changes (nice) plus additional prose about channels

\usepackage{url,hyperref,lineno,microtype}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage[onehalfspacing]{setspace}
\usepackage{gensymb} % For \degree

@@

@@

@@

% Solution described means we now need to do background of eye movements:
In an attempt to build a model combining brain, realistic
biomechanics \e{and} sensory feedback, we sought to extend our
previous work modelling the oculomotor system by adding a virtual,
biomechanical eye model able to make physically realistic
movements. The rotational state of the eye would then determine how
visual features in the virtual world were projected back into the
brain model.
%
The existing
model \citep{cope_basal_2017} is already able to capture sensory input
and convert it into a neural signal, assumed to specify the target of
a \e{saccadic eye movement}; a fast movement of the eyes which directs
the fovea to a region of interest in the field of view.
%
The oculomotor system is an excellent candidate for modelling because
its movements can be specified with only three degrees of freedom,
making it one of the simplest neuro-muscular systems in the body.
%
It is nevertheless behaviourally interesting, as saccadic eye
movements reveal information about decision making at a subconscious
level
\citep{deubel_saccade_1996,reppert_modulation_2015,marcos_determining_2016}.
The modelling of the oculomotor system is served by a large body of
behavioural data describing
saccades \citep{tipper_reaching_2001,walker_effect_1997,casteau_effect_2012},
many anatomical studies of the neural substrates involved
\citep{meredith_intrinsic_1998,isa_intrinsic_2002,isa_exploring_2009}
and electrophysiological data linking these together
\citep{hepp_spatio-temporal_1983,dorris_neuronal_1997,mcpeek_competition_2003,vokoun_circuit_2011}.
Furthermore, in the context of building \e{behaving} systems, a
necessary part of any model for which the behaviour requires visual
attention and decision making is a realistic mechanism for gathering
visual information.  This is obvious from extrinsic considerations---a
subject must look at a scene to make decisions or navigate within
it. It also follows for \e{intrinsic} reasons. For
example, \cite{howard_hand_1997} showed that visual cues affect reach
trajectories and the same group later demonstrated that reaching
affects the saccadic system
\citep{tipper_reaching_2001} suggesting a close relationship between these
neural systems. Building a behaving oculomotor system will
therefore assist future computational neurobehavioural modelling
efforts that involve reaching.

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a review,
see \cite{munoz_commentary:_2002}. One pathway takes information from
the retina directly into the superficial layers of the superior
colliculus in the brainstem
\citep{sterling_receptive_1971,linden_massive_1983,wu_involvement_1994}.
Activity within the superior colliculus then excites neurons in the
pons, medulla and rostral mid-brain \citep{sparks_brainstem_2002}.
and finally the motor neurons which innervate the extraocular muscles
\citep{fuchs_firing_1970,sparks_brainstem_2002}. This direct pathway
is responsible for the low latency saccades called express saccades
\citep{schiller_effect_1987,edelman_activity_1996}.
%
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \citep{schall_neural_1999}. Activity build-up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\citep{stanton_frontal_1988-1} and is also processed by the basal ganglia,
which participates in the selection of the winning saccade end point
\citep{stanton_frontal_1988}.
%
Although both cortical and subcortical paths produce a saccade target
signal in the superior colliculus, it is also possible for animals to
make relatively normal saccades even after the colliculus has been
ablated
\citep{wurtz_activity_1972,aizawa_reversible_1998},
though express saccades are lost with collicular lesions
\citep{schiller_effect_1987}. This makes the superior colliculus
a perplexing structure, being both critically involved in saccade
target specification \citep{sparks_sensory_1987} and saccade dynamic control
\citep{waitzman_superior_1991,goossens_optimal_2012} and yet dispensible.
The `backup pathway' likely incorporates the oculomotor vermis and
fastigial oculomotor region of the cerebellum which are known to
participate in the specification, dynamics and adaptation of saccadic
eye movements \citep{kleine_saccade-related_2003,takagi_effects_1998}.

There is a long history of modelling the oculomotor system. For a
comprehensive review, see \cite{girard_brainstem_2005}. Models of
individual subsystems have been proposed for brainstem
\citep{robinson_oculomotor_1975,scudder_new_1988,gancarz_neural_1998},
cerebellum \citep{quaia_model_1999,dean_modelling_1995,dean_learning_1994}
and superior colliculus
\citep{arai_two-dimensional_1994,moren_mechanism_2013,marino_spatial_2012}.
More recently, combined models have also been developed incorporating
sensory input \citep{cope_basal_2017} and driving a second order
differential equation representing the eye
\citep{tabareau_geometry_2007,nguyen_saccade_2014,thurat_biomimetic_2015}.
None of these models has yet fully closed the loop to produce a
behaving system operating freely within its environment. We argue that
developing integrated, closed-loop models of behaving systems offers
insights into the operation of neural systems that are not available
from input-~or output-assumption models.

@@

To the \ccg~model, we added a brainstem model and a biomechanical
eye model. The rate-coded brainstem model was taken from the
literature~\cite{gancarz_neural_1998} as the best-of-breed saccadic
burst generator~\cite{girard_brainstem_2005}.  The biomechanical eye
was implemented using the biomechanical modelling framework OpenSim.
These will be described below, along with a review of the \ccg~model,
but first we will give a description of the co-ordinate systems and
the modelling framework used.

@@

We implemented a saccadic burst generator (SBG) based on the
connectivity outlined in \cite{gancarz_neural_1998}. The SBG network
for two of the model's six channels is shown in Fig.~\ref{sbg}. We use
the word `channel' to mean a set of populations of neurons which are
involved in actuating a single extraocular muscle). Activity from the
output layer of superior colliculus
%
(SC\_avg) is fed into each channel, which sums the activity it
receives and processes it in populations each of a single neural
element representing all the neurons in that population. Each channel
of the SBG functions to create the motor neuron activations that are
required to accelerate the eye in a particular direction, then hold
the eye in its new position against the returning force generated by
the elastic properties of the muscles.  The required motor neuron
activations are therefore a combination of features: a brief burst of
increased activity that accelerates the eye; followed by a period of
activity that is less than the burst firing rate but higher than the
tonic rate that exists when the eye is at the centre. This holds the
eye in its new position.

@@

The generation of the `tonic' phase of the required time series is
achieved simply by integrating the bump over time and multiplying by a
some small gain factor.  This is the function of the tonic neurons
(TNs).  The firing rate of the motor neuron defines the amount of
force applied to the eye by that muscle.  Thus, the integral of the `bump'
defines how far the eye moves in that channel's direction.  The gain
and delay parameters in the LLBN-EBN-IBN-LLBN feedback loop therefore
have to be tuned such that the endpoint of the saccade is reasonably
accurate.  Furthermore the restoring force generated by the elasticity
of the muscles is dependent on the radial distance.  The value of the
new tonic firing rate, after the `bump' is dependent on the end
location of the eye.  If the ratio between the EBN firing rate and the
TN firing rate is not exactly correct, the eye will drift away from
the saccade endpoint after the saccade has been completed.  The EBN-TN
connection strength is therefore tuned such that the TN firing rate
yields a stable eye position across a range of eye eccentricities.

@@

We found the best parameters for the exponential in
Eq.~\ref{eq:weightmaps} ($i$ and $j$) by a manual tuning
process. After selecting values for $i$ and $j$ in either the
horizontal or vertical/oblique channels, we ran the model 6 times at
each of 8 target eccentricities (7\dg--14\dg) which were purely in the
direction of the newly parameterised channel. The training saccades
were produced as described below in Sect.~\ref{sec:singlesaccades},
with the same fixation and target luminances (crosses of magnitude 0.2
and 0.3) but with the fixation offset and target onset occurring at
0.2~s. We measured the end-point of the saccade by detecting the
location at which the saccade velocity had dropped below 0.005 of its
peak. We iterated until the mean saccade endpoint plotted versus
target was close to the ideal straight line---see
Fig.~\ref{sacc_vs_targ}(a) \& (b). We applied the same parameters to
both directions of each channel; $i_{up} = i_{down} = 0.00195$,
$j_{up} = j_{down} = 0.075$, $i_{left} = i_{right} = 0.0016$ and
$j_{left} = j_{right} = 0.067$. The resulting weight maps (where the
oblique maps are 1/10$^{\mathrm{th}}$ of the vertical maps, as
described earlier) are shown in Fig.~\ref{weightmaps}.
% From map_modelling2.m:
%    yh=0.0016*exp(0.067.*x);
%    yv=0.00195*exp(0.075.*x);

>>>>

\usepackage{url,hyperref,lineno,microtype}
\usepackage[english]{babel}
\usepackage{amsmath}
\usepackage[onehalfspacing]{setspace}
\usepackage{gensymb} % For \degree
\usepackage[usenames, dvipsnames]{color}

@@

% Use colour to indicate changes made following review. Can be
% set to this 'black' colour prior to final publication:
\newcommand{\black}{\textcolor{black}}
% Colour for reviewer 1 changes:
\newcommand{\rvwcolour}{\textcolor{OrangeRed}}
% Colour for reviewer 2 changes:
\newcommand{\rvwcolourtwo}{\textcolor{Cerulean}}
% Commands to use in the text. Change 'rvwcolour/rvwcolourtwo' to
% 'black' to hide review changes.
\newcommand{\rvw}[1]{\rvwcolour{#1}}
\newcommand{\rvwt}[1]{\rvwcolourtwo{#1}}

@@

% Remove after review:
Note: Changes based on Reviewer 1's comments \rvw{are in this colour}. For
Reviewer 2 \rvwt{we use this colour}.

@@

% Solution described means we now need to do background of eye movements:
In an attempt to build a model combining brain, realistic
biomechanics \e{and} sensory feedback, we sought to extend our
\rvw{previous work modelling the oculomotor system by adding a virtual,
biomechanical eye model able to make physically realistic
movements. The rotational state of the eye would then determine how
visual features in the virtual world were projected back into the
brain model.}
%
The existing
model \citep{cope_basal_2017} is already able to capture sensory input
and convert it into a neural signal, assumed to specify the target of
a \e{saccadic eye movement}; a fast movement of the eyes which directs
the fovea to a region of interest in the field of view.
%
The oculomotor system is an excellent candidate for modelling because
its movements can be specified with only three degrees of freedom,
making it one of the simplest neuro-muscular systems in the body.
%
It is nevertheless behaviourally interesting, as saccadic eye
movements reveal information about decision making at a subconscious
level
\citep{deubel_saccade_1996,reppert_modulation_2015,marcos_determining_2016}.
The modelling of the oculomotor system is served by a large body of
behavioural data describing
saccades \citep{tipper_reaching_2001,walker_effect_1997,casteau_effect_2012},
many anatomical studies of the neural substrates involved
\citep{meredith_intrinsic_1998,isa_intrinsic_2002,isa_exploring_2009}
and electrophysiological data linking these together
\citep{hepp_spatio-temporal_1983,dorris_neuronal_1997,mcpeek_competition_2003,vokoun_circuit_2011}.
Furthermore, in the context of building \e{behaving} systems, a
necessary part of any model for which the behaviour requires visual
attention and decision making is a realistic mechanism for gathering
visual information.  This is obvious from extrinsic considerations---a
subject must look at a scene to make decisions or navigate within
it. It also follows for \e{intrinsic} reasons. For
example, \cite{howard_hand_1997} showed that visual cues affect reach
trajectories and the same group later demonstrated that reaching
affects the saccadic system
\citep{tipper_reaching_2001} suggesting a close relationship between these
neural systems. \rvw{Building a behaving oculomotor system will
therefore assist future computational neurobehavioural modelling
efforts that involve reaching.}

Many neural populations are involved in the coding of saccadic eye
movements, only a very brief overview is given here; for a review,
see \cite{munoz_commentary:_2002}. One pathway takes information from
the retina directly into the superficial layers of the superior
colliculus in the brainstem
\citep{sterling_receptive_1971,linden_massive_1983,wu_involvement_1994}.
Activity within the superior colliculus then excites neurons in the
pons, medulla and rostral mid-brain \citep{sparks_brainstem_2002}.
and \rvw{finally the motor neurons} which innervate the extraocular muscles
\citep{fuchs_firing_1970,sparks_brainstem_2002}. This direct pathway
is responsible for the low latency saccades called express saccades
\citep{schiller_effect_1987,edelman_activity_1996}.
%
Information from the retina is also processed by visual cortex which
feeds through to the frontal eye fields in which activity is related
to reflexive and voluntary
saccades \citep{schall_neural_1999}. Activity build-up in the frontal
eye fields is transferred to the intermediate layers of the superior
colliculus
\citep{stanton_frontal_1988-1} and is also processed by the basal ganglia,
which participates in the selection of the winning saccade end point
\citep{stanton_frontal_1988}.
%
Although both cortical and subcortical paths produce a saccade target
signal in the superior colliculus, it is also possible for animals to
make relatively normal saccades even after the colliculus has been
ablated
\citep{wurtz_activity_1972,aizawa_reversible_1998},
though express saccades are lost with collicular lesions
\citep{schiller_effect_1987}. This makes the superior colliculus
a perplexing structure, being both critically involved in saccade
target specification \citep{sparks_sensory_1987} and saccade dynamic control
\citep{waitzman_superior_1991,goossens_optimal_2012} and yet dispensible.
The `backup pathway' likely incorporates the oculomotor vermis and
fastigial oculomotor region of the cerebellum which are known to
participate in the specification, dynamics and adaptation of saccadic
eye movements \citep{kleine_saccade-related_2003,takagi_effects_1998}.

There is a long history of modelling the oculomotor system. For a
comprehensive review, see \cite{girard_brainstem_2005}. Models of
individual \rvw{subsystems} have been proposed for brainstem
\citep{robinson_oculomotor_1975,scudder_new_1988,gancarz_neural_1998},
cerebellum \citep{quaia_model_1999,dean_modelling_1995,dean_learning_1994}
and superior colliculus
\citep{arai_two-dimensional_1994,moren_mechanism_2013,marino_spatial_2012}.
More recently, combined models have also been developed incorporating
sensory input \citep{cope_basal_2017} and driving \rvw{a second order
differential equation} representing the eye
\citep{tabareau_geometry_2007,nguyen_saccade_2014,thurat_biomimetic_2015}.
None of these models has yet fully closed the loop to produce a
behaving system operating freely within its environment. We argue that
developing integrated, closed-loop models of behaving systems offers
insights into the operation of neural systems that are not available
from input-~or output-assumption models.

@@

To the \ccg~model, we added a brainstem model and a biomechanical eye
model. \rvw{The rate-coded brainstem model was taken from the
literature}~\cite{gancarz_neural_1998} as the best-of-breed saccadic
burst generator~\cite{girard_brainstem_2005}.  The biomechanical eye
was implemented using the biomechanical modelling framework OpenSim.
These will be described below, \rvwt{along with a review of
the \ccg~model}, but first we will give a description of the
co-ordinate systems and the modelling framework used.

@@

We implemented a saccadic burst generator (SBG) based on the
connectivity outlined in \cite{gancarz_neural_1998}. The SBG network
for two of the model's six channels is shown in
Fig.~\ref{sbg}. \rvw{We use the word `channel' to mean a set of
populations of neurons which are involved in actuating a single
extraocular muscle. SBG channels are arranged in pairs, actuating
opposing muscles. There is one pair of channels which actuates the
superior and inferior rectus muscles, causing vertical rotations of
the eye in a roughly parasaggital plane (the eye moves up or
down). Another pair actuates the lateral and medial rectus muscles,
causing horizontal rotations of the eye. The third pair actuates the
superior and inferior oblique muscles which contribute to vertical as
well as oblique rotations.} Activity from the output layer of superior
colliculus
%
(SC\_avg) is fed into each channel, which sums the activity it
receives and processes it in populations each of a single neural
element representing all the neurons in that population. Each channel
of the SBG functions to create the motor neuron activations that are
required to accelerate the eye in a particular direction, then hold
the eye in its new position against the returning force generated by
the elastic properties of the muscles.  The required motor neuron
activations are therefore a combination of features: a brief burst of
increased activity that accelerates the eye; followed by a period of
activity that is less than the burst firing rate but higher than the
tonic rate that exists when the eye is at the centre. This holds the
eye in its new position.

@@

The generation of the `tonic' phase of the required time series is
achieved simply by integrating the bump over time and multiplying by a
some small gain factor.  This is the function of the tonic neurons
(TNs).  The firing rate of the motor neuron defines the amount of
force applied to the eye by that muscle.  Thus, \rvw{the integral}
of the `bump'
defines how far the eye moves in that channel's direction.  The gain
and delay parameters in the LLBN-EBN-IBN-LLBN feedback loop therefore
have to be tuned such that the endpoint of the saccade is reasonably
accurate.  Furthermore the restoring force generated by the elasticity
of the muscles is dependent on the radial distance.  The value of the
new tonic firing rate, after the `bump' is dependent on the end
location of the eye.  If the ratio between the EBN firing rate and the
TN firing rate is not exactly correct, the eye will drift away from
the saccade endpoint after the saccade has been completed.  The EBN-TN
connection strength is therefore tuned such that the TN firing rate
yields a stable eye position across a range of eye eccentricities.

@@

We found the best parameters for the exponential in
Eq.~\ref{eq:weightmaps} ($i$ and $j$) by a manual tuning
process. After selecting values for $i$ and $j$ in either the
horizontal or vertical/oblique channels, we ran the model 6 times at
each of 8 target eccentricities (7\dg--14\dg) which were purely in the
direction of the newly parameterised channel. The training saccades
were produced as described below in Sect.~\ref{sec:singlesaccades},
with the same fixation and target luminances (crosses of magnitude 0.2
and 0.3) but with the fixation offset and target onset occurring at
0.2~s. We measured the end-point of the saccade by detecting the
location at which the saccade velocity had dropped below 0.005 of its
peak. We iterated until the mean saccade endpoint plotted versus
target was close to the ideal straight line---see
Fig.~\ref{sacc_vs_targ}(a) \& \rvw{(b)}. We applied the same parameters to
both directions of each channel; $i_{up} = i_{down} = 0.00195$,
$j_{up} = j_{down} = 0.075$, $i_{left} = i_{right} = 0.0016$ and
$j_{left} = j_{right} = 0.067$. The resulting weight maps (where the
oblique maps are 1/10$^{\mathrm{th}}$ of the vertical maps, as
described earlier) are shown in Fig.~\ref{weightmaps}.
% From map_modelling2.m:
%    yh=0.0016*exp(0.067.*x);
%    yv=0.00195*exp(0.075.*x);
